---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{xkcd}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  warning = FALSE,
  message = FALSE
)
```

`xkcd` provides plot types for line charts, bar charts, pie charts, scatter plots and radar charts.

## Plot types

```{r setup}
library(xkcd)
library(dplyr)

df <- gapminder::gapminder |>
  filter(country %in% c("United States", "United Kingdom", "Canada", "Australia"))
```


### Line graph

```{r}
df |>
  x_chart(title = "my xkcd line graph") |>
  x_line(x = "year", y = "lifeExp", color = "country")
```


### Scatterplot

```{r}
iris |>
  x_chart() |>
  x_point(x = "Sepal.Width", y = "Sepal.Length", fill = "Species", size = 1.5) |>
  x_options(title = "scatterplot", palette = c("#f7feae", "#46aea0", "#045275"))
```


The `line` parameter in `x_point` controls wheather points are connected, this can be used like `geom_point + geom_line` in ggplot2.

```{r}
data.frame(x = 1:10, y = rnorm(10)) |>
  x_chart() |>
  x_point(x = "x", y = "y", line = TRUE) |>
  x_options(legend = FALSE)
```
### Bar and pie plot

```{r}
data.frame(
  character = c("Ross", "Rachel", "Joey", "Phoebe", "Monica", "Chandler"),
  relationships = c(16, 14, 17, 16, 13, 9)
) |>
  x_chart(title = "# of relationships in Friends") |>
  x_bar(x = "character", y = "relationships") |>
  x_options(legend = FALSE)
```

Stacked bar chart

```{r}
t <- as.data.frame(Titanic)
t <- t[rep(1:nrow(t), t$Freq), c("Class", "Survived")]

t |>
  count(Class, Survived) |>
  x_chart(title = "Titanic survivors by class and age") |>
  x_bar(x = "Class", y = "n", fill = "Survived") |>
  x_options(ylabel = "")
```

pie chart (does not support grouping)

```{r}
data.frame(
  x = 1:6,
  y = 1:6
) |>
  x_chart(title = "What I am made of") |>
  x_pie(x = "x", y = "y", radius = 0) |>
  x_options(legend_position = 2)
```


### Radar chart

```{r}
data.frame(
  x = rep(c("strength", "patience", "skills", "resilient", "flexibility"), 2),
  y = abs(rnorm(10)),
  g = rep(c("tiger", "lion"), each = 5)
) |>
  x_chart(title = "Tigher and lion: a random comparison") |>
  x_radar(x = "x", y = "y", color = "g")
```


## Configuration options

Common options that apply to all types of plots (set via `x_options`):

- `title`: plot title
- `xlabel`: x-axis title, default to x varibale
- `ylabel`: y-axis title, default to x variable
- `palette`: color palette for grouping variables, the number of colors should equal the number of groups
- `font_family`: fonts to use, default to "xkcd"
- `legend`: whether to show the legend
- `legend_position`: legend position, 1 (top-right), 2 (top-left), 3 (bottom-left), 4 (bottom-right)
- `time_format`: time format to use if the x values are time, chart.xkcd use \code{dayjs} to format time,  find the all the available formats at https://github.com/iamkun/dayjs/blob/dev/docs/en/API-reference.md#list-of-all-available-formats
- `x_breaks`: number of breaks for x axis, default is 3
- `y_breaks`: number of breaks for y axis, default is 3

specific options for each plot type:

- scatterplot
  - `size`:  dot size, default to 1
  - `line`: if dots should be connected, default to FALSE
- pie chart:
  - `radius`: arc radius, default to 0.5, for traditional pie chart set it to 0
- radar chart:
  - `size`: dot size, default to 1
  - `label`: whether to show labels for each dimension, default to TRUE
  - `ticks` number of ticks in each dimension